### 2024-09-06 manuscript.tex
Fix grammar

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between year
2010 and 2021 (Fig. \ref{fig:map}, Table \ref{tab:voyages} and
\ref{tab:voyage-references}). Altogether, the voyages and the station comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
smaller due to gaps in measurements. Missing days in the ceilometer data were
HMNZSW16 (7 days): 24--27 November, 10 December, 16--17 December 2016; MARCUS
(3 days): 8, 10 November, 10 December; MICRE (9 days): 7--8, 29 June, 5, 16
July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1 day):
24 January.

@@

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyage, measuring pressure, temperature, relative humidity and
global navigation satellite system coordinates. Derived thermodynamic (virtual
potential temperature, lifting condensation level, \ldots) and dynamic
quantities (wind speed and direction) for the measured vertical profiles were
calculated with rstool \citep{rstool}. Surface meteorological quantities were
measured continuously by an onboard automatic weather station or individual
instruments.

@@

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
by the nextGEMS project \citep{hohenegger2023}. nextGEMS has so far produced
four cycles of model runs. In our research, we used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b}. We could
not use Cycle 4 due to unavailability of the necessary variables. The
horizontal resolution of ngc3028 is about 5 km.  Unlike current general
circulation models (GCMs), the strom-resolving version of ICON does not use
convective and cloud parameterisation but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterisations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterised in this model. In our analysis, we used
four years of coupled simulation output between years 2021 and 2024
(inclusive). Because the model is free-running, weather and climate
oscillations (such as the El Niño--Southern Oscillation) are not expected to be
equivalent to reality at the same time and place. To compare with observations
taken during different years (2010--2021, inclusive), we compared the model
output with observations at the same time of year and geographical location.

@@

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator. It performs radiative transfer calculations to derive equivalent
lidar backscattered radiation in an atmospheric model, which can then
be compared with observed lidar backscatter \citep{kuma2021}. For this purpose,
it takes atmospheric fields of cloud fraction, liquid and ice mass mixing ratio,
temperature and pressure fields as an input and can be run offline (on the
model output rather than inside the model code). The lidar simulator in the
ALCF is based on the instrument simulator Cloud Feedback Model
Intercomparison Project (CFMIP) Observation Simulator Package (COSP)
\citep{bodas-salcedo2011}. After the backscattered radiation is calculated, a
cloud mask, cloud occurrence by height and the total cloud fraction are
determined. ALCF has already been used by several research teams for model and
reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.
Absolute calibration of the observed backscatter was performed by comparing
the measured clear-sky molecular backscatter statistically with simulated
clear-sky molecular backscatter.

@@

The Vaisala CL51 and CL31 are ceilometers operating at a near-infrared
wavelength of 910 nm. The maximum range 15.4 km (CL51), 7.7 km (CL31 with 5 m
vertical resolution) and 7.550 (CL31 with 10 m vertical resolution).
The vertical resolution is 10 m (5 m configurable). The wavelength is affected
by water vapour absorption of about 20\% in the mid-latitudes
\citep{wiegner2015,wiegner2019}.  The instrument data files containing
uncalibrated backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF to
produce a cloud mask and cloud occurrence by height and the total cloud
fraction. CL51 can also be configured to emulate CL31.

For the MARCUS voyage, Vaisala CL31 were made available by the Atmospheric
Radiation Measurement (ARM) processed to a maximum range of 7.545 km, vertical
resolution of 30 m and 252 vertical levels.

@@

The Lufft CHM 15k is a ceilometer operating at a near-infrared wavelength of
1064 nm. The maximum range is 15.4 km, vertical resolution is 5 m in the near
range (up to 150 m) and 15 m above, the sampling rate is 2 s, and the number of
vertical levels is 1024.  NetCDF files containing uncalibrated backscatter
produced by the instrument were processed with the ALCF to produce a cloud mask
and cloud occurrence by height and the total cloud fraction.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric
model \citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° in latitude and 0.625° in longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72, which we
interpolate to height levels as needed. In our analysis we use the following
products: 1-hourly instantaneous 2D single-level diagnostics (M2I1NXASM) for
2-m temperature and humidity; 3-hourly instantaneous 3D assimilated
meteorological fields (M2I3NVASM) for cloud quantities, pressure and
temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation and sea ice; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

We partitioned our data into two subsets by cyclonic activity. For this
purpose, we used a cyclone tracking algorithm to identify extratropical and
polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data, We
used the open source cylone tracking package CyTRACK \citep{perez-alarcon2024}.
Generally, what constitutes an EC is considered relatively arbitrary due to a
very large variability of ECs \citep{neu2013}.  We used the mean sea level
pressure field and horizontal wind speed fields as input to the CyTRACK
algorithm. The algorithm uses pressure and wind speed thresholds as well as
tracking across time steps to identify cyclone centres and radii. From this
information, we could classify geographical areas as either cyclonic or
non-cyclonic. Due to a relatively small total area covered, we chose a circle
of a double radius (relative to one identified by CyTRACK) centred at the
cyclone centre as a cyclonic area for every time step and cyclone. All other
areas were identified as non-cyclonic. For identifying cyclones in the
observations and the reanalyses, ERA5 pressure and wind fields were used as the
input to CyTRACK.  This is justified by the fact that the large-scale pressure
and wind fields in ERA5 are likely sufficiently close to the reality. For
identifying cyclones in ICON, its own pressure and wind fields were used as the
input to CyTRACK, because the model is free-running, and thus the pressure and
wind fields are different from reality.

@@

We partitioned our data into two subset by stability. We determined this by
calculating lower tropospheric stability (LTS) as the difference between
potential temperature at 700 hPa and the surface.  Based on a histogram of LTS
in ERA5 and MERRA-2 calculated at all voyage tracks and stations (Fig.
\ref{fig:lts}), we determined a dividing threshold of 12 K for relatively
unstable (< 12 K) and relatively stable (>= 12 K) conditions.

@@

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence and type. We
used this dataset to train a convolutional artiﬁcial neural network (ANN) of
the U-Net type \citep{ronneberger2015} to recognise profiles with precipitation
from lidar backscatter. Samples of short time intervals (10 min) of
near-surface backscatter (0–250 m) were classified as clear, rain, snow and
fog, using the synoptic observations as a training dataset. From these, a
binary classification of profiles as wet (rain or fog) and dry (clear or fog)
was derived.

@@

The radiation calculations were done in such a way that they always represent
averages of daily means. This is done in order to be consistent with the CERES
SYN1deg data, which are expressed as daily means. Therefore, for every
instantaneous profile in the simulated lidar data is assigned a daily mean
radiation value corresonding to the day (in the Coordinated Universal Time;
UTC). In turn, the average radiation during the entire voyage or station
measurement is calculated as the average of the profile values. In the observed
lidar data, the daily mean radiation value is taken from the spatially and
temporally co-located CERES SYN1deg data for the day (in UTC). The voyage or
station measurement average is calculated in the same way.

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{List of voyages and stations. Start, end and days (UTC; inclusive) indicate the time period when the vessel was south of 40°S. Abbreviations: ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA), Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the nearest integer. CL51/31 indicates CL51 configured to emulate CL31.}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL31    & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & [Macquarie Is. station]       & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON and two reanalyses MERRA-2 and ERA5 (Fig.
\ref{fig:cloud-occurrence-panel}). We aggregated the sources by calculating
the averages and percentiles of all individual profiles (Fig.
\ref{fig:cloud-occurrence}). The analysis shows that the total cloud fraction
(determined as the fraction of proﬁles with clouds at any height in the lidar
cloud mask) is underestimated in ICON and reanalyses by about 10\% and 20\%,
respectively.

@@

The observations show cloud occurrence peaking at the surface, whereas models
show higher peak (at about 500 m). The models underestimate total cloud
fraction by 10--20\%, and show a strong decrease in cloud occurrence near the
surface, but this is not happening in the observations. ICON and ERA5
overestimate cloud occurrence at the peak (between 0--1 km). Above 1 km, ICON
and MERRA-2 understimate cloud occurrence, but ERA5 is very accurate. The too
high peak in models is partly supported by the LCL distribution, which peaks
higher in the models then in the observations (at the surface), although this
is not very pronounced. The northern zone shows stronger peak of cloud
occurrence near the surface than the southern zone, and this could be because
higher latitudes have more unstable atmospheric profiles. The northern and
southern zones show similar biases in models as in the general case, but ERA5
does not overestimate the peak in the northern zone (near-surface cloud
occurrence is still strongly understimated). Cyclonic situations have a larger
amount observed cloudiness, including the peak and total cloud fraction. In
these situations, the models are doing a relatively good job of getting the
vertical profile of cloud occurrence right, but still tend to underestimate
cloud occurrence above 1 km and near the surface. Non-cyclonic situations are
similar to the general case. In stable stituations cloud occurrence peaks
strongly at the surface in observations, compared to unstable situations, where
the peak is more more obtuse and spread across 0--1 km. In stable situations
the models are doing a fairly good job, but overestimate cloud occurrence at
the peak below 1 km, above 1 km the show similar biases as in the general case.
In unstable situations the bias in ICON is very pronounced, with a much
stronger peak at about 500 m, ERA5 is underestimating cloud occurrence below 1
km (especially near the surface), and MERRA-2 even more strongly. In all
situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in unstable and non-cyclonic
conditition. It also shows peak of cloud occurrence at higher altitude than obs
(500 m vs. near the surface) and correspondingly its LCL tends to be also
higher. MERRA-2 can be generally characterised as underestimating cloud
occurrence at nearly all altitudes as well as the total cloud fraction, but
mostly above and below 500 m (the peak at 500 m is well-represented). It
struggles the most in the northern zone, and in unstable situations. ERA5 can
be generally characterised as representing cloud occurrence correctly above
about 1--1.5 km, overestimating below, but underestimating near-surface cloud
occurrence (0--500 m). The total cloud fraction is strongly understimated in
all situations. It has tendency towards understimation in the northern zone and
unstable stiuations, and conversely overestimating in the southern zone and
stable conditions.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and
stations for the observed (OBS) and simulated lidar cloud mask. The total cloud
fraction (CF), average shortwave (SW) and longwave (LW) are shown in the
legend, and the relative frequency of occurrence (RFO) of the subset. The bands
span the 16$^\mathrm{th}$--84$^\mathrm{th}$ percentile calculated from the set
of all voyages and stations. The subsets \textbf{(a--g)} are described in the
text (Section \ref{sec:results}).
}
\label{fig:cloud-occurrence}
\end{figure}

>>>>

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, resupply
vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B.
Palmer}, Her Majesty's New Zealand Ship (HMNZS) \emph{Wellington} and one
sub-Antarctic station (Macquarie Island) in the SO south of 40°S between 2010
and 2021 (Fig. \ref{fig:map}, Table \ref{tab:voyages} and
\ref{tab:voyage-references}). Altogether, the voyages and the station comprised
2421 days of data south of 40°S, but the availability of ceilometer data was
smaller due to gaps in measurements. Missing days in the ceilometer data were
HMNZSW16 (7 days): 24--27 November, 10 December, 16--17 December 2016; MARCUS
(3 days): 8, 10 November, 10 December; MICRE (9 days): 7--8, 29 June, 5, 16
July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1 day):
24 January.

@@

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages, measuring pressure, temperature, relative humidity, and
global navigation satellite system coordinates. Derived thermodynamic (virtual
potential temperature, lifting condensation level, etc.) and dynamic quantities
(wind speed and direction) for the measured vertical profiles were calculated
with rstool \citep{rstool}. Surface meteorological quantities were measured
continuously by an onboard automatic weather station or individual instruments.

@@

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
by the nextGEMS project \citep{hohenegger2023}. nextGEMS has so far produced
four cycles of model runs. In our research, we used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b}. We could
not use Cycle 4 due to the unavailability of the necessary variables. The
horizontal resolution of ngc3028 is about 5 km.  Unlike current general
circulation models (GCMs), the strom-resolving version of ICON does not use
convective and cloud parameterisation but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterisations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterised in this model. In our analysis, we used
four years of coupled simulation output between years 2021 and 2024
(inclusive). Because the model is free-running, weather and climate
oscillations (such as the El Niño--Southern Oscillation) are not expected to be
equivalent to reality at the same time and place. To compare with observations
taken during different years (2010--2021, inclusive), we compared the model
output with observations at the same time of year and geographical location.

@@

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator. It performs radiative transfer calculations to derive equivalent
lidar backscattered radiation in an atmospheric model, which can then be
compared with observed lidar backscatter \citep{kuma2021}. For this purpose, it
takes atmospheric fields of cloud fraction, liquid and ice mass mixing ratio,
temperature, and pressure fields as an input and can be run offline (on the
model output rather than inside the model code). The lidar simulator in the
ALCF is based on the instrument simulator Cloud Feedback Model Intercomparison
Project (CFMIP) Observation Simulator Package (COSP) \citep{bodas-salcedo2011}.
After the backscattered radiation is calculated, a cloud mask, cloud occurrence
by height, and the total cloud fraction are determined. ALCF has already been
used by several research teams for model and reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.
Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter.

@@

The Vaisala CL51 (Fig. \ref{fig:map}d) and CL31 are ceilometers operating at a
near-infrared wavelength of 910 nm. The maximum range is 15.4 km (CL51), 7.7 km
(CL31 with 5 m vertical resolution), and 7.550 km (CL31 with 10 m vertical
resolution).  The vertical resolution is 10 m (5 m configurable). The
wavelength is affected by water vapour absorption of about 20\% in the
mid-latitudes \citep{wiegner2015,wiegner2019}.  The instrument data files
containing uncalibrated backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF to
produce a cloud mask and cloud occurrence by height and the total cloud
fraction. CL51 can also be configured to emulate CL31.

For the MARCUS voyage, Vaisala CL31 was made available by the Atmospheric
Radiation Measurement (ARM) processed to a maximum range of 7.545 km, a
vertical resolution of 30 m, and 252 vertical levels.

@@

The Lufft CHM 15k (Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
rate is 2 s, and the number of vertical levels is 1024.  NetCDF files
containing uncalibrated backscatter produced by the instrument were processed
with the ALCF to produce a cloud mask and cloud occurrence by height and the
total cloud fraction.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° in latitude and 0.625° in longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72, which we
interpolate to height levels as needed. In our analysis we use the following
products: 1-hourly instantaneous 2D single-level diagnostics (M2I1NXASM) for
2-m temperature and humidity; 3-hourly instantaneous 3D assimilated
meteorological fields (M2I3NVASM) for cloud quantities, pressure, and
temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation and sea ice; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

We partitioned our data into two subsets by cyclonic activity. For this
purpose, we used a cyclone tracking algorithm to identify extratropical and
polar cyclones (ECs and PCs) over the SO in the reanalysis and ICON data. We
used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii. From this information, we could classify geographical areas as
either cyclonic or non-cyclonic. Due to a relatively small total area covered,
we chose a circle of a double radius (relative to one identified by CyTRACK)
centred at the cyclone centre as a cyclonic area for every time step and
cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality.

@@

We partitioned our data into two subsets by stability. We determined this by
calculating lower tropospheric stability (LTS) as the difference between the
potential temperature at 700 hPa and the surface.  Based on a histogram of LTS
in ERA5 and MERRA-2 calculated at all voyage tracks and stations (Fig.
\ref{fig:lts}), we determined a dividing threshold of 12 K for relatively
unstable (< 12 K) and relatively stable (>= 12 K) conditions.

@@

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence and type. We
used this dataset to train a convolutional artiﬁcial neural network (ANN) of
the U-Net type \citep{ronneberger2015} to recognise profiles with precipitation
from lidar backscatter. Samples of short time intervals (10 min) of
near-surface backscatter (0–250 m) were classified as clear, rain, snow, and
fog, using the synoptic observations as a training dataset. From these, a
binary classification of profiles as wet (rain or fog) and dry (clear or fog)
was derived.

@@

The radiation calculations were done in such a way that they always represent
averages of daily means. This is done in order to be consistent with the CERES
SYN1deg data, which are expressed as daily means. Therefore, for every
instantaneous profile in the simulated lidar data is assigned a daily mean
radiation value corresponding to the day (in the Coordinated Universal Time;
UTC). In turn, the average radiation during the entire voyage or station
measurement is calculated as the average of the profile values. In the observed
lidar data, the daily mean radiation value is taken from the spatially and
temporally co-located CERES SYN1deg data for the day (in UTC). The voyage or
station measurement average is calculated in the same way.

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{
List of voyages and stations. Start, end, and days (UTC; inclusive)
indicate the time period when the vessel was south of 40°S. Abbreviations:
ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA),
Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the
nearest integer. CL51/31 indicates CL51 configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL31    & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & [Macquarie Is. station]       & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, and two reanalyses, MERRA-2 and ERA5 (Fig.
\ref{fig:cloud-occurrence-panel}). We aggregated the sources by calculating the
averages and percentiles of all individual profiles (Fig.
\ref{fig:cloud-occurrence}). The analysis shows that the total cloud fraction
(determined as the fraction of proﬁles with clouds at any height in the lidar
cloud mask) is underestimated in ICON and reanalyses by about 10\% and 20\%,
respectively.

@@

The observations show cloud occurrence peaking at the surface, whereas models
show a higher peak (at about 500 m). The models underestimate total cloud
fraction by 10--20\% and show a strong decrease in cloud occurrence near the
surface, but this is not happening in the observations. ICON and ERA5
overestimate cloud occurrence at the peak (between 0--1 km). Above 1 km, ICON
and MERRA-2 understimate cloud occurrence, but ERA5 is very accurate. The too
high peak in models is partly supported by the LCL distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced. The northern zone shows a stronger peak of cloud
occurrence near the surface than the southern zone, and this could be because
higher latitudes have more unstable atmospheric profiles. The northern and
southern zones show similar biases in models as in the general case, but ERA5
does not overestimate the peak in the northern zone (near-surface cloud
occurrence is still strongly understimated). Cyclonic situations have a larger
amount of observed cloudiness, including the peak and total cloud fraction. In
these situations, the models are doing a relatively good job of getting the
vertical profile of cloud occurrence right, but still tend to underestimate
cloud occurrence above 1 km and near the surface. Non-cyclonic situations are
similar to the general case. In stable situations cloud occurrence peaks
strongly at the surface in observations, compared to unstable situations, where
the peak is more obtuse and spread across 0--1 km. In stable situations, the
models are doing a fairly good job, but overestimate cloud occurrence at the
peak below 1 km; above 1 km, they show similar biases as in the general case.
In unstable situations the bias in ICON is very pronounced, with a much
stronger peak at about 500 m, ERA5 is underestimating cloud occurrence below 1
km (especially near the surface), and MERRA-2 even more strongly. In all
situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in unstable and non-cyclonic
condititions. It also shows a peak of cloud occurrence at higher altitude than
obs (500 m vs. near the surface), and correspondingly, its LCL tends to be also
higher. MERRA-2 can be generally characterised as underestimating cloud
occurrence at nearly all altitudes as well as the total cloud fraction, but
mostly above and below 500 m (the peak at 500 m is well-represented). It
struggles the most in the northern zone and in unstable situations. ERA5 can be
generally characterised as representing cloud occurrence correctly above about
1--1.5 km, overestimating below, but underestimating near-surface cloud
occurrence (0--500 m). The total cloud fraction is strongly understimated in
all situations. It has a tendency towards understimation in the northern zone
and unstable situations, conversely, overestimating in the southern zone and
stable conditions.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and
stations for the observed (OBS) and simulated lidar cloud mask. The total cloud
fraction (CF), average shortwave (SW) and longwave (LW) are shown in the
legend, and the relative frequency of occurrence (RFO) of the subset is shown
below.  The bands span the 16$^\mathrm{th}$--84$^\mathrm{th}$ percentile
calculated from the set of all voyages and stations. The subsets
\textbf{(a--g)} are described in the text (Section \ref{sec:results}).
}
\label{fig:cloud-occurrence}
\end{figure}
